---
title: "HW5_Balasubramanian_Venkataraman"
author: "Venkataraman Balasubramanian"
date: "3/8/2022"
output: html_document


knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Shiny Dashboard to Visualize IPL Cricket Data

For the shiny dashboard assignment I will be using the same dataset that I used for the table assignment - The IPL cricket data. The goal of this dashboard is to get the users familiarize with the teams along with the number of wins they had over the course of IPL plus how each team fared every year.  

The shiny dashboard will comprise of two pages - 

- Analysis by Year - Enable the user to search by Year and see how each teams performed (the number of wins is a criterion to perform this comparison)  
- Analysis by Teams - Enable the user to search by teams and see how the team performed over the course of IPL  

The dashboard will also be interactive. The users can search by year or teams and the charts in the dashboard pages will display data accordingly per the search criteria   

There is one visuals in the "Analysis by Year" page -

 - A horizontal bar chart showing number of wins by all the teams in the selected year  

There are two visuals in the "Analysis by Teams" page -

 - A vertical bar chart showing the number of games won by the selected team over the course of IPL  
 - A table showing the details of the dames won by the selected team over the course of IPL  
 
 I hope my cohort will use the shiny dashboard to get themselves familiarized not only with the game of cricket but also the teams that play IPL & the history of IPL as well  
 
## Load Packages and data  
```{r warning=FALSE, message=FALSE}
library(shiny)
library(shinydashboard)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(plotly)
library(DT)
library(kableExtra)

#Load IPL cricket data
data = read.csv("IPL_Game_Metadata.txt", sep = "|")
data$GameYear = format(as.Date(data$GameDate, format = "%Y-%m-%d"), "%Y%")
```

## Shiny Dashboard Code
```{r}

server = function(input, output){
  
  ## Bar Chart for Winners
  output$bar_chart_winners = renderPlotly({
  
  in_year = input$in_year
  #Filter out rows with blank winners
  data_winners = data %>%
  filter(Winner != "")

  #Filter out the temporary teams
  data_winners = data_winners %>%
  filter(Winner %in% c("Sunrisers Hyderabad", "Royal Challengers Bangalore", "Chennai Super Kings", "Rajasthan Royals", "Mumbai Indians", "Kolkata Knight Riders", "Kings XI Punjab", "Delhi Daredevils", "Delhi Capitals", "Deccan Chargers")) %>% filter(GameYear == in_year)

  bar_chart_winners = ggplot(data = data_winners, mapping = aes(x = Winner)) +
  geom_bar(mapping = aes(fill = Winner)) +
  coord_flip() +
  theme_classic()+
  scale_fill_brewer(palette = "BrBG") +
  xlab("Number of Wins") +
  ylab("Team Name") 
  
  ggplotly(bar_chart_winners)
  })
  
  ## Bar chart for games played every year
  output$bar_chart_game_count = renderPlotly({
    
  in_name = input$in_name
  
  data_game_count = data %>% filter(GameYear <= 2018 & Winner == in_name)
  
  bar_chart_game_count = ggplot(data = data_game_count, mapping = aes(GameYear))+
  geom_bar(mapping = aes(fill = GameYear)) +
  theme_classic()+
  scale_fill_brewer(palette = "Spectral") +
  xlab("Year") +
  ylab("Number of Games")+
  labs(fill = "Year")

    ggplotly(bar_chart_game_count)
  })
  
  ## Data Table
  output$tbl_data = renderDataTable({
    
    in_name = input$in_name
    data_winners %>% select(GameYear,HomeTeam,AwayTeam,Winner) %>% filter(Winner==in_name)
  })
  
  output$in_name1 = renderText({
    input$in_name
  })
  
    output$in_name2 = renderText({
    input$in_year
  })
}

ui = dashboardPage(
  skin = "blue",
  dashboardHeader(
    title = "IPL"
  ),
  dashboardSidebar(
    sidebarMenu(
      textInput("in_name", "Team Name", value = "Chennai Super Kings"),
      textInput("in_year","Year", value = "2008"),
      menuItem("Analysis By Teams", tabName = "by_teams"),
      menuItem("Analysis By Year", tabName = "by_year")

    )
  ),
  dashboardBody(
    #First Page
    tabItems(
      tabItem("by_teams",
              h2("Games Won by ", textOutput("in_name1", inline = TRUE), "in IPL"),
              h3("Number of Games Won"),
                box(plotlyOutput("bar_chart_game_count"), width = 500),
              h3("Details of Games Won"),
                box(dataTableOutput("tbl_data"), width = 500)
              ),

    #Second Page
    tabItem("by_year",
            h2("Games Won by Teams in Year ", textOutput("in_name2", inline = TRUE)),
              box(plotlyOutput("bar_chart_winners"), width = 500)
    )
    )
  )
)

shinyApp(ui,server)
```

